---
title: "Q5_Powerpoint_Present"
author: "Precious Nhamo"
date: "`r Sys.Date()`"
output: ioslides_presentation
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("Health.R", encoding = "UTF-8")
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Source your code file from the same folder
source("Health.R", encoding = "UTF-8")
# Load required packages efficiently with pacman 
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, broom, kableExtra)
# Load necessary packages here too, if not inside the script
```


```

# Introduction

Analysis of key determinants of health care based on WHO-funded data

Focus on sleep, stress, physical activity, and demographic differences

Goal: Provide practical insights for improving health outcomes

## Sleep as a Proxy for Health
How does high stress impact the benefits of adequate sleep?

Analysis using boxplots and regression models with interaction effects

To what extent can stress diminish or negate sleep’s positive health effects?

## Role of Physical Activity
How does physical activity influence the relationship between sleep and health?

Can physical inactivity offset the benefits of good sleep quality?

## Demographic Differences
Which age groups are most vulnerable to poor sleep, high stress, and inactivity?

How do lifestyle differences between younger and older adults affect these health determinants?

# Speaker Notes
In this presentation, I will highlight how sleep serves as a proxy for overall health, emphasizing the complex role that high stress plays in potentially reducing the positive effects of good sleep. We will also explore how physical activity modifies this relationship, examining whether inactivity can cancel out the benefits of quality sleep. Finally, I will discuss which demographic groups—particularly younger versus older adults—are most impacted by poor sleep, stress, and inactivity, providing actionable insights for targeted health interventions.


## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3




## Sleep stress interaction box and whisker and regression results 
```{r , echo=FALSE}
library(broom)
library(kableExtra)

# Tidy the model summary with confidence intervals and significance stars
tidy_model <- broom::tidy(model_sleep_stress_controls) %>%
  mutate(
    signif = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      p.value < 0.1   ~ ".",
      TRUE           ~ ""
    ),
    Estimate = sprintf("%.3f", estimate),
    `Std. Error` = sprintf("%.3f", std.error),
    `t value` = sprintf("%.3f", statistic),
    `Pr(>|t|)` = sprintf("%.4f", p.value),
    term = gsub("`", "", term)  # clean variable names if needed
  ) %>%
  select(term, Estimate, `Std. Error`, `t value`, `Pr(>|t|)`, signif)

# Print with kableExtra
tidy_model %>%
  kable(
    caption = "Table: Regression Results for Weight Change on Sleep, Stress, and Controls",
    col.names = c("Variable", "Estimate", "Std. Error", "t value", "P-value", ""),
    align = c("l", "r", "r", "r", "r", "c"),
    booktabs = TRUE
  ) %>%
  kable_styling(latex_options = c("hold_position", "striped"), font_size = 10)

```

```{r , echo = FALSE}
plot(boxplot_sleep_stress)
```
