---
title: "Data Science Exam"

author: 
  - name: "Precious Nhamo"

abstract: |
  This document comprises the responses to **Questions 1 to 5** of the 2025 Data Science examination,
  along with separate analyses (e.g., a PowerPoint presentation) completed in accordance with the exam
  instructions and organised within the designated folder.

bibliography: "r-references.bib"

floatsintext: yes
linenumbers: no
draft: no
mask: no

figurelist: no
tablelist: no
footnotelist: no

classoption: "man"
output: papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```



# Question1
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

## Results

##Discussion

#Question2
## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

##Results

##Discussion

#Question3
## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.
##Results

##Discussion

#Question4
## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.
##Results

##Discussion

#Question5
## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.
##Results

##Discussion









\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
