# Question 5: Health Data Analysis — Weight Change, Sleep, Stress & Physical Activity

This question investigates how stress levels, sleep quality, and lifestyle factors affect weight change across age groups. It uses regression models, stratified summaries, and insightful visualizations to identify key health risk patterns.

---

## Project Structure

├── Data/
│ └── HealthCare.csv
├── Results/
│ ├── Age_Distribution.png
│ ├── Age_Summary.csv
│ ├── Gender_Distribution.csv
│ ├── Weight_Change_Summary.csv
│ ├── Caloric_Activity_Summary.csv
│ ├── Sleep_Stress_Boxplot.png
│ ├── Sleep_Risk_Pivot.csv
│ ├── Stress_Risk_Pivot.csv
│ ├── Regression_Models.txt
│ └── OnAir.txt
├── health_analysis_script.R
└── README.md

##Notes 
Age groupings were categorised from teen , young adults , early career adults , pre and post pension (senior) to properly seperate effects during different stages of life .
 Age < 20 ~ "Teenagers (<20)",
      Age >= 20 & Age < 30 ~ "Adults (20-29)",
      Age >= 30 & Age < 40 ~ "Early Career (30-39)",
      Age >= 40 & Age < 50 ~ "Middle-aged (40-49)",
      Age >= 50 & Age < 65 ~ "Preretirement (50-64)",
      Age >= 65 ~ "Seniors (65+)",
      
Risk flagings were created to flag high stress , poor sleep and sedentary life style 
Pivot tables show stress / sleep impact across age groups 
We used regression models for examining weight changes 
Box and whisker of weight change across sleep-stress combinations 
And a summary of the on air script is provided.

## Requirements

Install and load the following packages using `pacman`:

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, skimr, kableExtra, broom, GGally, readr, future, furrr, modelsummary)
```

```{r}
source("C:/Users/pmnha/my-new-project/22660348/Question5/Code/Health.R")
main_analysis("C:/Users/pmnha/my-new-project/22660348/Question5/Data/HealthCare.csv")
```

* **High Stress Risk**:
  Individuals with a `Stress Level` of **7 or higher** are classified as "Yes" for high stress risk, meaning they are considered to have high stress. Those with stress levels below 7 are classified as "No".

* **Poor Sleep Risk**:
  People whose `Sleep Quality` is either **"Poor"** or **"Fair"** are labeled as "Yes" for poor sleep risk. Those with better sleep quality (e.g., "Good" or "Excellent") are labeled "No".

* **Weight Change Risk**:
  This is a bit more complex:

  * If someone gained **more than 5 pounds** and their `Daily Caloric Surplus/Deficit` is **greater than 300 calories** (i.e., consistently eating more), they are considered "Yes" for weight change risk.
  * If someone lost **more than 5 pounds** and their caloric deficit is **less than -500 calories** (eating significantly less), they are also "Yes" for risk.
  * Otherwise, "No".

* **Sedentary Lifestyle**:
  Those who report their `Physical Activity Level` as exactly **"Sedentary"** are classified as "Yes" for sedentary lifestyle, otherwise "No".

* **Stress Group**:
  A simplified grouping where individuals with stress level ≥7 are tagged as **"High Stress"**, and the rest as **"Low Stress"**.

* **Sleep Group**:
  Similarly, people with "Poor" or "Fair" sleep are grouped as **"Poor Sleep"**, and others as **"Good Sleep"**.

* **Combined Group**:
  This is a combination label created by pasting together the sleep group and stress group, e.g., **"Poor Sleep + High Stress"**, **"Good Sleep + Low Stress"**, etc. This allows for analysis of combined sleep-stress categories.

* **Factors for Modeling**:
  Some variables are converted to factors (categorical variables) for modeling purposes:

  * `sleep_risk` (Poor Sleep / Good Sleep)
  * `stress_risk` (High Stress / Low Stress)
  * `physical_activity` (levels from the original physical activity data)
  * `age_group` (age categories created elsewhere)





